---
title: "R Usage"
---

## Overview

SCG is implemented in Python but can be accessed from R via the [reticulate](https://rstudio.github.io/reticulate/) package. This page demonstrates how to load the pre-computed SCG results from the [data application](data-application.qmd) and work with them in R. We also show how to fit the model from scratch in R for users who prefer that workflow.

## Setup

::: {.callout-note}
Replace the path in `use_python()` with the path to your own Python installation that has `scr` installed. To find it, run `which python` or `which python3` in your terminal. Please see [reticulate](https://rstudio.github.io/reticulate/) for further guidance. 
:::

```{r}
library(reticulate)
library(ggplot2)
library(dplyr)

use_python("~/.pyenv/versions/3.11.14/bin/python")
np <- import("numpy")
```

## Loading Pre-computed Results

The results from the data application are saved in `../results/` and can be loaded directly in R without re-fitting the model.

### SCG Classifications

```{r}
results <- read.csv("../results/results.csv")
spot_info <- read.csv("../data/spot_info.csv")
gene_names <- read.csv("../data/gene_names.csv")

scgs <- results[results$scg == "True", ]
cat(nrow(scgs), "SCGs identified out of", nrow(results), "edges\n")
```

### Correlation Fields

```{r}
Corr_est <- np$load("../results/Corr_est.npy")
dim(Corr_est)
```

## Visualizing Spatially Varying Edges

```{r}
boundary_colors <- c("T" = "#7651A6", "N" = "#2E7D6E", "I" = "#C4952D")

plot_corr_field <- function(
  Corr_est,
  spot_info,
  g,
  g_prime,
  gene_names = NULL
) {
  field <- Corr_est[, g + 1, g_prime + 1] # convert 0-indexed to 1-indexed

  df <- data.frame(
    x = spot_info$x,
    y = spot_info$y,
    corr = field,
    region = spot_info$region
  )

  vmax <- max(abs(df$corr))

  g_label <- if (!is.null(gene_names)) {
    gene_names$gene[g + 1]
  } else {
    as.character(g)
  }
  g_prime_label <- if (!is.null(gene_names)) {
    gene_names$gene[g_prime + 1]
  } else {
    as.character(g_prime)
  }

  ggplot(df, aes(x = x, y = y, color = corr)) +
    geom_point(size = 2.0) +
    scale_color_gradient2(
      low = "#2166ac",
      mid = "white",
      high = "#d6604d",
      midpoint = 0,
      limits = c(-vmax, vmax),
      name = "Correlation"
    ) +
    coord_equal() +
    ggtitle(paste0(g_label, "—", g_prime_label)) +
    theme_void() +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, size = 11)
    )
}

plot_corr_field(
  Corr_est,
  spot_info,
  g = 2,
  g_prime = 4,
  gene_names = gene_names
)
plot_corr_field(
  Corr_est,
  spot_info,
  g = 39,
  g_prime = 49,
  gene_names = gene_names
)
```

## Fitting the Model from R

Users who wish to fit SCG from scratch in R can do so via `reticulate`. The workflow mirrors the Python data application exactly.

```{r}
#| eval: false
scr <- import("scr")
torch <- import("torch")

Y_np <- as.matrix(read.csv("../data/Y.csv"))
Y_np <- t(Y_np)

dims <- dim(Y_np)
N <- dims[1]
P <- dims[2]
K <- L <- as.integer(round(2 * log(P)))

init <- scr$scr_init(
  Y = Y_np,
  L = L,
  K = K,
  a_delta = c(2.1, 3.1),
  b_delta = 1.0,
  a0 = 5.0,
  b0 = 0.5,
  nu = 5.0,
  spca_center = TRUE,
  spca_alpha = 0.01,
  spca_ridge_alpha = 0.01,
  device = scr$DEVICE,
  dtype = scr$DTYPE
)

kernel_hypers <- scr$pick_kernel_params(
  init = init,
  S = as.matrix(spot_info[, c("x", "y")])
)
K_np <- scr$rq_kernel(
  as.matrix(spot_info[, c("x", "y")]),
  v2 = 1.0,
  rho = 0.1 * kernel_hypers$rho,
  alpha = kernel_hypers$alpha
)

fit <- scr$cavi(
  torch$as_tensor(Y_np, device = scr$DEVICE, dtype = scr$DTYPE),
  torch$as_tensor(K_np, device = scr$DEVICE, dtype = scr$DTYPE),
  init,
  max_iter = 60L,
  verbose = FALSE
)

results <- scr$classify_edges(fit$params, M = 500L, alpha = 0.1)
```

::: {.callout-note}
When passing data from R to Python via reticulate, NumPy arrays are non-writable by default. This triggers a PyTorch warning during `torch$as_tensor` that can be safely ignored.
:::

## Session Info

```{r}
sessionInfo()
```
